---
title: "Challenge 1"
author: "Pablo Barbera"
date: "January 21, 2016"
output: html_document
---

The purpose of this challenge is to analyze the Twitter network of the Members of the U.S. Congress.

In the `sna` folder of the GitHub repository, you will find two data files:

- `congress-twitter-network-edges.csv` contains the edges of this network. Note that unlike in the previous example, these nodes are now directed: they indicate whether the legislator in the `source` column follows the legislator in the `target` column.
- `congress-twitter-network-nodes.csv` contains information about each of the nodes. The only important variables you need to use for this challenge are: `id_str` (the unique Twitter ID for each legislator; same as in the edge list), `name` (full name of each legislator), `party` (Republican, Democrat or Independent), and `chamber` (`rep` for the House of Representatives, `sen` for the Senate).

The first step will be to read these two datasets into R and construct the igrahp object. How many nodes and edges does this network have?

```{r}
nodes <- read.csv("congress-twitter-network-nodes.csv")
edges <- read.csv("congress-twitter-network-edges.csv")

...
```

This network is too large for us to visualize it directly with R, so let's try to learn more about it using what we have learned so far.

How many components does this network have? As you will see, in this particular case it makes sense that we work only with the giant component.

```{r}

```

Who are the most relevant Members of Congress, according to different measures of centrality? Note that this is a directed network, which means there is a difference between indegree and outdegree.

```{r}

```

What communities can you find in the network? Use the additional node-level variables to try to identify whether these communities overlap with any of these other attributes. Try different community detection algorithms to see if you get different answers.

```{r}

```

Finally, we'll try to visualize the network as well. Instead of plotting it in the Viewer window, we'll write directly to a PDF file. I have added a few options here for you so that it's faster, but note that this will probably take 1-2 minutes.

```{r}
set.seed(777)
fr <- layout_with_fr(g, niter=1000)
V(g)$color <- ifelse(V(g)$party=="Republican", "red", "blue") # clue
V(g)$shape <- ifelse(V(g)$chamber=="sen", "square", "circle") # clue
V(g)$label <- NA
V(g)$size <- authority_score(g)$vector * 5

pdf("congress-network.pdf")
par(mar=c(0,0,0,0))
plot(g, edge.curved=.25, edge.width=.05, edge.arrow.mode=0)
dev.off()
```




