---
title: "Social network analysis with R: modeling"
author: "Pablo Barbera"
date: "January 21, 2016"
output: html_document
---

`ergm` package works with object in `network` format (not `igraph`), so we need to convert first between packages using the `intergraph` package.

```{r, echo=FALSE, message=FALSE}
nodes <- read.csv("data/congress-twitter-network-nodes.csv")
edges <- read.csv("data/congress-twitter-network-edges.csv")
library(igraph)
g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
g <- decompose(g)[[1]]

library(intergraph)
library(network)
net <- asNetwork(g)

```

Edges = number of edges in the network. Similar to a constant. exp(edges) close to observed density.
Mutual = number of mutual edges in the network. If positive, it means reciprocity is more common than expected. If you see an edge from i to j; then you are exp(mutual) times more likely to see an edge from j to i as well.
nodematch = probability that an edge between nodes with the same value on this variable exists. exp(nodematch) times more likely for an edge to exist when values match than otherwise.

```{r}
#install.packages("ergm")
library(ergm)
reg1 <- ergm(net ~ edges + mutual)
summary(reg1) 
mcmc.diagnostics(reg1)
```

```{r}
reg2 <- ergm(net ~ edges + mutual + nodecov("followers_count") + nodefactor("chamber"))
summary(reg2)

reg3 <- ergm(net ~ edges + nodefactor("chamber") +
               absdiff("followers_count", pow=1/2) + nodematch("party") + 
               nodematch("chamber") + nodematch("gender"))
summary(reg3)
exp(coef(reg3))
```

