---
title: "Social network analysis with R: modeling"
author: "Pablo Barbera"
date: "January 21, 2016"
output: html_document
---

`ergm` package works with object in `network` format (not `igraph`), so we need to convert first between packages using the `intergraph` package.

```{r, echo=FALSE, message=FALSE}
nodes <- read.csv("congress-twitter-network-nodes.csv")
edges <- read.csv("congress-twitter-network-edges.csv")
library(igraph)
g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
g <- decompose(g)[[1]]

library(intergraph)
library(network)
net <- asNetwork(g)

```

Edges = number of edges in the network. Similar to a constant. exp(edges) close to observed density.
Mutual = number of mutual edges in the network. If positive, it means reciprocity is more common than expected. If you see an edge from i to j; then you are exp(mutual) times more likely to see an edge from j to i as well.
nodematch = probability that an edge between nodes with the same value on this variable exists. exp(nodematch) times more likely for an edge to exist when values match than otherwise.

```{r}
#install.packages("ergm")
library(ergm)
reg1 <- ergm(net ~ edges + mutual)
summary(reg1) 
mcmc.diagnostics(reg1)
```

```{r}
reg2 <- ergm(net ~ edges + mutual + nodecov("followers_count") + nodefactor("chamber"))
summary(reg2)

reg3 <- ergm(net ~ edges + nodefactor("chamber") +
               absdiff("followers_count", pow=1/2) + nodematch("party") + 
               nodematch("chamber") + nodematch("gender"))
summary(reg3)
exp(coef(reg3))
```

Latent space model for networks
```{r}
# install.packages("latentnet")
library(latentnet)
lat <- ergmm(net ~ euclidean(d=1), verbose=TRUE, 
             control=ergmm.control(sample.size = 100, burnin=100, interval = 1))

test <- prcomp(as.matrix(g[])*1)

library(ca)
res <- ca(as.matrix(g[])*1)

```


```{r}
#install.packages("NetData")
library(NetData)
data(studentnets.ergm173)
edges <- edges[edges$alter_id %in% nodes$std_id,]
g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
net <- asNetwork(g)

reg1 <- ergm(net ~ edges + mutual + nodematch("gnd") + nodematch("rce"))
summary(reg1)

```
