---
title: "Challenge 2"
author: "Pablo Barbera, Dan Cervone"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `taxi-trips.csv`. Note that a dictionary of the variables names for this data set is in the file `taxi-dictionary.pdf`, though we have split the pickup and dropoff "datetime" variables into "date" and "time" variables for each. 

### Sample solutions

Please note there are many possible (equally excellent) solutions! 

```{r}
taxi <- read.csv("./taxi-trips.csv", stringsAsFactors=FALSE)
```


1. Convert the `pickup_time` and `dropoff_time` variables into numeric variables (for instance, minutes past 12AM), so that you can create a new variable called `trip_time` (in minutes).

```{r}
split.pickup <- strsplit(taxi$pickup_time, ":")
split.dropoff <- strsplit(taxi$dropoff_time, ":")

pickup_min <- sapply(split.pickup, function(x) 60 * as.numeric(x[1]) + as.numeric(x[2]) + as.numeric(x[3]) / 60)
dropoff_min <- sapply(split.dropoff, function(x) 60 * as.numeric(x[1]) + as.numeric(x[2]) + as.numeric(x[3]) / 60)
taxi$trip_time <- dropoff_min - pickup_min

summary(taxi$trip_time)
summary(taxi$fare_amount)
# there are clearly some bad values, so let's remove them
taxi <- subset(taxi, trip_time > 0 & trip_time < 60 & fare_amount > 0)
```

2. NYC taxi fares start at 2.50, and then add 2.50 per mile and 0.50 per minute stopped or in traffic (though the data set includes some flat fare or negotiated fare rides). See if you can reverse engineer this information by running a regression of `fare_amount` on `trip_distance` and `trip_time`. 

```{r}
# regression with RateCodeID == 1 (standard fare)
fare.mod <- lm(fare_amount ~ trip_distance + trip_time, data=subset(taxi, RateCodeID == 1))
summary(fare.mod)
```

Unfortunately we can't quite recover the fare rules. This is probably because the time is only charged when the vehicle is stopped or moving very slowly, which we can't measure here. 

3. Using this regression model, see if you can identify passengers who have been overcharged.

We do this by identifying very large residuals.

```{r}
resids <- residuals(fare.mod)
which(resids > 20)
subset(taxi, RateCodeID == 1)[which(resids > 20), ]
```

4. What fare do you predict for a solo passenger traveling 2.1 miles in a trip lasting 14 minutes?

```{r}
predict(fare.mod, newdata=data.frame(trip_distance=2.1, trip_time=14))
```

5. Using logistic regression, identify factors (e.g. fare amount, pickup time, number of passengers, etc.) that are predictive of whether or not passengers pay by credit card.

```{r}
taxi$credit_card <- taxi$payment_type == 1 # binary variable for CC payment

split.pickup <- strsplit(taxi$pickup_time, ":")
taxi$pickup_hour <- sapply(split.pickup, function(x) x[1])
cc.mod <- glm(credit_card ~ passenger_count + fare_amount + pickup_hour, data=taxi, family=binomial(logit))
summary(cc.mod)
```

We see that, holding all else constant, credit card payments are associated with fewer numbers of passengers and higher fare amounts. We also see CC payments are more likely late at night (hours "20" to "03"). CC payments are least likely early in the morning (hours "05" to "10").
