---
title: "RMarkdown setup"
author: "Pablo Barbera"
date: "January 20, 2016"
output: html_document
---

Welcome to the course! 

The goal of this first lesson is to provide an introduction to R for non-programmers. We will provide an overview of the fundamentals of this programming language, how to write modular code, and best practices for using R for data analysis.

__Why R?__ 

- Currently used by most statisticians and social scientists interested in data analysis; and also becoming one of most [popular languages in Data Science](http://www.kdnuggets.com/2015/05/r-vs-python-data-science.html).
- Open source: makes it highly customizable and easily extensible through ["packages"](https://cran.r-project.org/web/packages/) (over 7,500 and counting!).
- Powerful tool to generate elegant and effective plots, both with built-in functions and additional packages such as lattice or ggplot2.
- Command-line interface and scripts favors reproducibility.
- Excellent documentation and online help resources.

We will be using [RStudio](https://www.rstudio.com/) to interact with R, and write our annotated R code using [Markdown](http://rmarkdown.rstudio.com).

__RStudio__ is an open-source integrated development environment (IDE). The main advantage of RStudio with respect to other graphical interfaces, such as R GUI (the default), is that it integrates a powerful built-in text editor as well as other tools for plotting, debugging, and workspace management.

__Markdown__ is a simple formatting syntax to generate HTML or PDF documents. In combination with R, it will generate a document that includes the comments, the R code, and the output of running such code.

You can embed R code in chunks like this one:

```{r}
1 + 1
```

You can run each chunk of code one by one, by highlighting the code and clicking `Run` (or pressing `Ctrl + Enter` in Windows or `command + enter` in OS X). You can see the output of the code in the console right below, inside the RStudio window.

Alternatively, you can generate (or __knit__) an html document with all the code, comment, and output in the entire `.Rmd` file by clicking on `Knit HTML`.

You can also embed plots, for example:

```{r}
x <- c(1, 3, 4, 5)
y <- c(2, 6, 8, 10)
plot(x, y)
```

If you run the chunk of code, the plot will be generated on the panel on the bottom right corner. If instead you knit the entire file, the plot will appear after you view the html document.

Using R + Markdown has several advantages: it leaves an "audit trail" of your work, including documentation explaining the steps you made, it allows you to easily correct errors (you can always fix them and run the script again), and after you have finished, you can generate a PDF or HTML version of your work.

We will be learning more about R Markdown over the next few modules. For more details on how to use it, see <http://rmarkdown.rstudio.com>.