---
title: "First steps in R"
author: "Pablo Barbera"
date: "January 20, 2016"
output: html_document
---

#### Using R as a calculator

The most basic functionality of R is using it as a calculator

```{r}
10 / 2
sqrt(100) + sqrt(9)
exp(1)
```

What makes R very powerful is that you can store results as "objects"

```{r}
x = 5
y = 10
```

If you look at the `Environment` panel on the top right, you can see that these numbers are stored in memory.

Then you can do operations with them, the same way you would do with numbers:

```{r}
x * y
```

You can also save combinations of objects as new objects

```{r}
z <- x * y
```

Note the `<-` sign. That's the assignment operator. You can also use `=`, although `<-` is generally preferred (we can discuss why later).

To print one of these objects in the console, you can do:

```{r}
print(z)
```

`print` here is a function: it takes a series of arguments (in this case, the object `z`) and returns a value (`50`).

This is equivalent to just typing the name of the object in the console. (What's going on behind the scenes is that R is calling the default function to print this object; which in this case is just `print`).

```{r}
z
```


#### Vectors and basic visualization

If objects have more than one element, they are called "vectors"

```{r}
height = c(181, 175, 179, 175, 176)
```

Note the use of the `c` (concatenate) function to put all 5 numbers together.
For example, this vector contains the height in cm of five people.

```{r}
print(height)
```

You can create as many vectors as you want

```{r}
weight = c(80, 61, 74, 58, 67)
```

To visualize the relationship between the two variables, you can use the `plot` function

```{r}
plot(x=height, y=weight)
```

`plot` is a function just like `print`, but it takes multiple arguments. First, you specify the variable you want in the x-axis: `x=height`. Then, the one on the y-axis: `y=weight`.

#### Finding help

Now imagine that you want to add a title to your plot, but don't know how. You can use the help page for `plot`:
```{r}
?plot
```

And find that "main" is the option you're looking for

```{r}
plot(x=height, y=weight, main="Relationship between height and weight")
```

This is just a basic plot -- we will cover more advanced examples this afternoon

#### Installing and loading packages

There are many packages that expand the basic functionalities of R. For example, imagine that we want to make a map of the U.S. The package "maps" is exactly what we're looking for.

First, we install it (you only need to do this once) typing the following:

```{r, echo=FALSE}
install.packages("maps", repos = "http://cran.rstudio.com/")
```

Alternatively:
1) Go to the "Packages" tab on the bottom right panel
2) Click on "Install Packages"
3) Type "maps" and click on "Install"

Now, we load the package (you will need to do this every time you want to use a function from the package)
```{r}
library(maps)
```

And we can use the "map" function to  draw a map of the US. Note that we need to specify the option 'database="state"'

```{r}
map(database="state")
```

