---
title: "Introduction to data analysis"
author: "Pablo Barbera, Dan Cervone"
date: "January 20, 2016"
output: html_document
---

### Working with vectors


```{r}
math_scores <- c(72, 58, 85, 83, 87, 66, 72, 93, 87, 84)
reading_scores <- c(50, 56, 76, 79, 74, 57, 80, 98, 90, 78)
```

We can also generate vector with the `seq` function:

```{r}
seq(1, 10, by=1)
```

Note that this is equivalent to:

```{r}
1:10
```

Functions to determine the structure of vectors:

```{r}
length(math_scores)
head(math_scores)
tail(math_scores)
```

How to subset vectors:

```{r}
math_scores[5]
reading_scores[5]
math_scores[-1]
```

Using vectors to subset vectors:

```{r}
math_scores[c(1,2,3)]
math_scores[5:7]
math_scores[-(1:5)] # note parentheses!
```

`which` function to compare all values of a vector

```{r}
which(math_scores>=80)
high <- which(math_scores>=80)
math_scores[high]
```

Statistical functions:

```{r}
mean(math_scores)
sum(math_scores)
sd(math_scores)
min(math_scores)
max(math_scores)
```

All in one function:

```{r}
summary(math_scores)
```

Vectorized operations:
```{r}
math_scores > 80
math_scores + reading_scores
```

### Data frames

```{r}
df <- data.frame("math" = math_scores, "reading" = reading_scores)
```

Access variables:
```{r}
df$math
df$math[1:3]
df$math[-(1:5)] # note parentheses!
```

Access rows and columns. DF internally is a matrix
```{r}
df[1,]
df[,1]
```

Subset data frame in same way as vectors
```{r}
df[1:5,]
df[df$math > 85, ]
```

Identify the maximum
```{r}
best_student <- which.max(df$math)
df[best_student, ]
```

### Working with a data frame

```{r}
df <- read.csv("humansofnewyork.csv")
```

`View(df)`

```{r}
str(df)
```


```{r}
head(df)
head(df, n=1)
tail(df)

```

```{r}
df$message[100]
df[100,]
mean(df$likes_count)
max(df$likes_count)
```

```{r}
which(df$likes_count==3384088)
```

```{r}
df[1263,]
```

```{r}
df[which(df$likes_count==max(df$likes_count)),]
```
