---
title: "Challenge 1"
author: "Pablo Barbera, Dan Cervone"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `humansofnewyork.csv`. Read this data into `R` and answer the questions below, trying not to rely too much on the previous code we used!

```{r}
hony <- read.csv("./humansofnewyork.csv", stringsAsFactors=FALSE) # read in data, character vectors treated as character vectors (not factors)
str(hony) # examine its structure
```


1. How many status updates have been posted on this page?

```{r}
table(hony$type) # type of facebook post
sum(hony$type == "status")
```

2. What is the total number of likes, comments, and shares it received?

```{r}
total.likes <- sum(hony$likes_count)
total.comm <- sum(hony$comments_count)
total.shares <- sum(hony$shares_count)
total.likes + total.comm + total.shares # wow!
```

3. What is the content of the post with the highest number of shares?

```{r}
max(hony$shares_count) # maximum num shares
top.post <- which.max(hony$shares_count)
hony$message[top.post]
```

4. What was the date in which the first photo was posted?

The dates we have are characters, so we can't sort them. However, we can notice that the rows in our data frame are ordered by date. Thus let's find the first row that includes a photo.

```{r}
head(hony$created_time) # in order
tail(hony$created_time) # in order
first.photo <- min(which(hony$type == "photo"))
hony$created_time[first.photo] # October 1, 2011
```

5. What is the total number of likes that the page has ever received, excluding its most popular post?

```{r}
max.likes <- max(hony$likes_count) # likes on most popular page
sum(hony$likes_count) - max.likes
```

6. How many posts have received more than 1,000,000 likes?

```{r}
sum(hony$likes_count > 1000000)
```

7. What was the total number of shares received by posts published each year?

```{r}
year <- substr(hony$created_time, 1, 4) # extracts year from date created variable
tapply(hony$shares_count, year, sum) # sum of shares by year
```
