---
title: "Control Flow and Modularity"
author: "Pablo Barbera, Dan Cervone"
date: "January 20, 2016"
output: html_document
---

Our introduction to `R` has focused on data manipulation, for which `R` is uniquely well suited as a programming language. However, `R` does all the things essential to writing programs in other languages; data analysis often requires or is eased by incorporating these tasks.

### Loops

"For" loops are probably the most common type of loop and are easily implemented in `R`. For instance, we can get the first 40 terms in the Fibonacci sequence using a for loop.

```{r}
fib <- c(0, 1, rep(NA, 38)) # initialize fib sequence
for(i in 3:40) {
  fib[i] <- fib[i-1] + fib[i-2]
}
```

Note that most of the time that `sapply` can be used instead of a for loop, it will be faster. For instance we can create the `Month` variable for the Daily Show data using a for loop instead of `sapply`:

```{r}
daily <- read.csv("./daily_show_guests.csv") # double check working directory
show.split <- strsplit(as.vector(daily$Show), "/")
Month <- rep(NA, length(show.split)) # initialize Month variable
for(i in 1:length(Month)) {
  Month[i] <- show.split[[i]][1]
}

system.time(replicate(100, {
  Month <- rep(NA, length(show.split))
  for(i in 1:length(Month)) {
    Month[i] <- show.split[[i]][1]
  }
}))
system.time(replicate(100, {
  sapply(show.split, function(x) x[1])
}))
```

#### While loops

Another common loop type is a "while" loops, which repeats indefinitely until some stopping condition is met (or alternatively, *while* a looping condition remains true). Such a loop is very useful, for instance, for determining convergence of a numerical algorithm. To illustrate this, let's calculate the logarithm of 0.1 using the Taylor series expansion 

$$\log(x) = \sum_{n=1}^{\infty}\frac{(-1)^{n+1}}{n}(x - 1)^n$$

```{r}
x <- 0.1
n <- 1
abs.dif <- Inf
value <- 0
while(abs.dif > .000001 & n <= 1000) {
  new.value <- value + (((-1) ^ (n + 1)) * (x - 1) ^ n) / n
  abs.dif <- abs(new.value - value) # convergence check
  value <- new.value
  n <- n + 1
}
value
log(0.1)
n
```

### If statements

Depending on whether a condition is true or false, we might want to execute different chunks of code.

```{r}
compare_xy <- function(x, y) {
  if (x < y) {
    print("y is greater than x")
  } else if (x > y) {
    print("x is greater than y")
  } else {
    print("x and y are equal")
  }
}
compare_xy(3, 4)
compare_xy(4, 3)
compare_xy(1, 1)
```

### Modular code

When writing lots of code, it's a good idea to keep things as modular as possible. This means breaking complicated tasks into smaller, simpler tasks. You may benefit by using these small chunks in other sections of code you write, and they are easier to check and debug if you are gettin gerrors. 

For example, let's say we want to create a new Daily Show data frame with 3 variables for month, day, and year, as well as a variable for the number of previous appearances the guest had on the show. We could do:

```{r}
modify_daily <- function(daily) {
  # first convert "MM/DD/YY"" to month/day variables
  show.split <- strsplit(as.vector(daily$Show), "/")
  month <- sapply(show.split, function(x) x[1])
  day <- sapply(show.split, function(x) x[2])
  year <- sapply(show.split, function(x) x[3])
  
  # now get previous appearances
  prev <- rep(0, nrow(daily))
  for(i in 2:length(prev)) {
    prev[i] <- sum(daily$Raw_Guest_List[1:(i-1)] == daily$Raw_Guest_List[i])
  }
  return(data.frame(year, month, day, prev))
}

new_daily <- modify_daily(daily)
```

But an even better (more modular) approach is to write separate functions that generate the date variables and previous appearance variable:

```{r}
separate_MDY <- function(mdy, sep="/") {
  mdy.split <- strsplit(as.vector(mdy), sep)
  month <- sapply(mdy.split, function(x) x[1])
  day <- sapply(mdy.split, function(x) x[2])
  year <- sapply(mdy.split, function(x) x[3])
  return(data.frame(year, month, day))
}
prev_app <- function(vec) {
  prev <- rep(0, length(vec))
  for(i in 2:length(prev)) {
    prev[i] <- sum(vec[1:(i-1)] == vec[i])
  }
  return(prev)
}
modify_daily_2 <- function(daily) {
  mdy <- separate_MDY(daily$Show)
  prev <- prev_app(daily$Raw_Guest_List)
  return(data.frame(mdy, prev))
}

new_daily <- modify_daily_2(daily)
```

The advantage here is that we could use the smaller components `separate\_MDY` and `prev_app` in other sections of code later on, if only one of these two tasks were needed. Also, if we input a new data file that stores dates as "MM-DD-YY" instead of "MM/DD/YY", the generic `separate_MDY` module can handle this by changing supplying the `sep="-"` argument; we wouldn't need to change the main `modify\_daily\_2` function. 