---
title: "Challenge 2"
author: "Pablo Barbera, Dan Cervone"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `congress-terms.csv`. Please read the data into `R` and answer the questions below.

### Sample solutions

Please note there are many possible (equally excellent) solutions! 

```{r}
congress <- read.csv("./congress-terms.csv", stringsAsFactors=FALSE) # read in data
```

1. How many total congressmen have represented each political party?

```{r}
table(congress$party)
# there are clearly some bad rows here, so let's eliminate them
congress <- subset(congress, party %in% c("D", "I", "R")) # keep only dems, reps, inds
table(congress$party)
```

2. Who was the youngest congressman elected and in how many congresses did he/she serve?

```{r}
min(congress$age) # youngest age
youngest <- which.min(congress$age) # which row in the data has this age
first.youngest <- congress$firstname[youngest]
last.youngest <- congress$lastname[youngest]
paste(first.youngest, last.youngest)
sum(congress$firstname == first.youngest & congress$lastname == last.youngest) # total terms
```


3. In how many congresses did the House of Representatives have a Democratic majority?

```{r}
house <- subset(congress, chamber == "house")
house.dems <- tapply(house$party, house$congress, function(x) sum(x == "D"))
house.reps <- tapply(house$party, house$congress, function(x) sum(x == "R"))
sum(house.dems > house.reps)
```

4. Which state has had the oldest congressmen on average?

```{r}
congress$age <- as.numeric(congress$age) # convert age into numeric variable
age.table <- tapply(congress$age, congress$state, mean) # mean age by state
age.table[which.max(age.table)]
```

5. For which state is the proportion of total Democrats throughout history closest to the proportion of Republicans?

```{r}
total.dems <- tapply(congress$party, congress$state, function(x) sum(x == "D"))
total.reps <- tapply(congress$party, congress$state, function(x) sum(x == "R"))
prop.dems <- total.dems / (total.dems + total.reps)
abs.prop.diff <- abs(prop.dems - 0.5) # absolute difference in proportion of dems/reps
abs.prop.diff[which.min(abs.prop.diff)] # North Dakota!
```

6. How many times has the majority party in the House of Representatives changed?

```{r}
dem.majority <- house.dems > house.reps # logical vector if dems have majority
party.change <- 0 # variable to count party changes
for(i in 2:length(dem.majority)) {  # this will loop through each congress
  if(dem.majority[i] != dem.majority[i-1]) { # check if the majority changed from previous congress
    party.change<- party.change + 1 # if so, add 1 to party.change count
  }
}
party.change
```

7. Which state in the 113th congress has the most experienced congressmen (measured my the average number of congresses in which they served)?

```{r}
experience <- rep(0, nrow(congress)) # new experience variable
for(i in 2:nrow(congress)) {
  # count previous appearances of each congressman
  experience[i] <- sum(congress$firstname[1:(i-1)] == congress$firstname[i] & congress$lastname[1:(i-1)] == congress$lastname[i])
}
congress$experience <- experience # add this to congress data frame
congress.113 <- subset(congress, congress == 113)
exp.table <- tapply(congress$experience, congress$state, mean)
exp.table[which.max(exp.table)]
```
