---
title: "Challenge 2"
author: "Pablo Barbera, Dan Cervone"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `congress-terms.csv`. Try not to look at the previous code we used!

```{r}

```


1. How many total congressmen have represented each political party?

```{r}

```

2. Who was the youngest congressman elected and in how many congresses did he/she serve?

```{r}

```


3. In how many congresses did the House of Representatives have a Democratic majority?

```{r}

```

4. Which state has had the oldest congressmen on average?

```{r}

```

5. For which state is the proportion of total Democrats throughout history closest to the proportion of Republicans?

```{r}

```

6. How many times has the majority party in the House of Representatives changed?

```{r}

```

7. Which state in the 113th congress has the most experienced congressmen (measured my the average number of congresses in which they served)?

```{r}

```
