---
title: "Introduction to R text analysis"
author: Alex Hanna
date: January 21, 2016
output: html_document
---

[&laquo; Text Analysis Module](../text/README.md)

In this section we are going to focus on some basic string manipulation in R. We will be using some of the R base functionaity as well as the `stringr` package. For more information on the `stringr` package, check out the [documentation vignette](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html).

R stores the basic string in a character vector. `length` gets the number of items in the vector, while `nchar` is the number of characters in the vector.

```{r results = "hide"}
str1 <- 'This is a string.'
is.character(str1)
length(str1)
nchar(str1)
```

We can also have multiple strings within one vector. A vector of strings is the most simple kind of data structure for storage.

```{r results = "hide"}
v_str1 <- c('This is a string.', 'This is also a string.', 'A third one goes here.')
length(v_str1)
nchar(v_str1)
sum(nchar(v_str1))
```

It can also be part of a larger data frame of information. For this, load the dataset of tweets from the bullying project. Use `head` to look at the first few lines. 

```{r results = "hide"}
df.tweets <- read.csv("bullying.csv", header = TRUE, stringsAsFactors = FALSE)
is.character(df.tweets$text)
head(df.tweets$text)
```

Now we can start to dig into the text a little bit more. Given the construction of the dataset, we can expect that there will be many instances of the word "bully." We can use the `grep` command to identify these. `grep` returns the index where the word occurs.

```{r results = "hide"}
grep('bully', 'That guy is a buly.')
grep('bully', 'That guy is a bully. Fixed that for you.')
grep('bully', c('That guy is a buly.', 'That guy is a bully. Fixed that for you.', 'How dare you correct me.'))
```

Within the dataset, we can use the results of `grep` to get particular rows. First, check if the number of tweets mentioning "bully" match the total number of rows in the data frame.
```{r results = "hide"}
nrow(df.tweets)
length(grep('bully', df.tweets$text))
```

That didn't seem to work. That's because matching is case-sensitive. Use `grepl` to match to a lowercase version.
```{r results = "hide"}
length(grepl('bully', df.tweets$text))
```

<!--
	1. String vectors
	2. Basic operations
		a. grep
		b. str_to_lower, str_to_upper, str_to_title
		c. str_match
		d. substr
		e. str_replace

-->
