---
title: "Challenge 1"
author: Alex Hanna, Pablo Barbera, Dan Cervone
date: January 21, 2016
output: html_document
---

[&laquo; Text Analysis Module](../text/README.md)

Write commands that help you answer the following questions about the bullying dataset.

1. How many of these tweets are in English? (there is a column that can help you here)
```{r}
df.tweets <- read.csv("bullying.csv", header = TRUE, stringsAsFactors = FALSE)
table(df.tweets$lang)
table(df.tweets$lang)["en"]
```

2. How many of the tweets contain a link? How many mention another Twitter user?
```{r}
length(grep("HTTPLINK", df.tweets$text))
length(grep("USER", df.tweets$text))
```

3. How many tweets contain a hashtag?
```{r}
length(grep("#", df.tweets$text))
```

4. How many of the tweets mention anything about a school? (include any word you think relate to school)
```{r}
length(grep("school|institute|scholar|class", df.tweets$text, ignore.case=TRUE))
```

5. Display the first 20 tweets which mention a school.
```{r}
df.tweets$school <- grepl("school|institute|scholar|class", df.tweets$text, ignore.case=TRUE)
school.tweets <- df.tweets[df.tweets$school==TRUE,]
head(school.tweets[,c("text")])
```

6. How many tweets mention "mom" and "dad"?
```{r}
df.tweets$mom <- grepl("mom", df.tweets$text, ignore.case=TRUE)
df.tweets$dad <- grepl("dad", df.tweets$text, ignore.case=TRUE)
table(df.tweets$mom==TRUE & df.tweets$dad == TRUE)
```

7. Display the first 15 tweets which mention "mom" and "dad". Note any peculiar findings.
```{r}
head(df.tweets$text[df.tweets$mom==TRUE & df.tweets$dad==TRUE], n=15)
```

